---
title: "Setting Up Git and GitHub for R Users"
author: "Your Name"
date: today
format: gfm
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE)
```

# Git and GitHub Setup for R Users

This tutorial will guide you through setting up Git and GitHub for use with
Positron. By following these steps, you'll be able to version control your R
projects and collaborate with others through GitHub.

## Prerequisites

Before starting, ensure you have:

* Positron installed (latest version recommended)
* R installed (latest version recommended)
* Internet connection to download packages and access GitHub

## Installation Process

### 1. Sign up for GitHub

If you don't already have a GitHub account, you'll need to create one:

1. Go to [GitHub's join page](https://github.com/join)
2. Follow the prompts to create your account
3. Verify your email address

GitHub is a web-based hosting service for Git repositories, offering
collaboration features, issue tracking, and more.

### 2. Install Git

Git is a version control system that tracks changes in your code over time.
Install Git based on your operating system:

**For Windows users:**

```{r}
#| label: install-git-windows
#| eval: false
# Download and install from:
browseURL("https://git-scm.com/download/win")


```

**For Mac users:**

```{r}
#| label: install-git-mac
#| eval: false
# Download and install from:
browseURL("https://git-scm.com/download/mac")
```

After installation, restart Positron to ensure it recognizes Git.

### 3. Configure Git with Positron

Now we'll configure Git to work with Positron by setting your user information
and authentication.

#### Set your user name and email

Use the `usethis` package to configure your Git username and email:

```{r}
#| label: config-git
#| eval: false
# Install usethis if needed
if (!requireNamespace("usethis", quietly = TRUE)) {
  install.packages("usethis")
}

# Set your Git username and email
# Replace with your actual name and email
usethis::use_git_config(user.name = "John Doe", user.email = "jdoe@mail.com")
```

Your Git commits will be associated with this name and email.

#### Create a Personal Access Token (PAT)

GitHub requires a Personal Access Token for secure authentication:

```{r}
#| label: create-token
#| eval: false
# This will open GitHub in your browser to create a token
usethis::create_github_token()
```

Follow the prompts on GitHub to generate your token. Make sure to:

1. Give it a descriptive name (e.g., "Positron-Access")
2. Set an appropriate expiration date
3. Select scopes (permissions) - typically "repo" and "workflow" are sufficient
4. Copy the token immediately after creation (you won't be able to see it again!)

#### Store your Personal Access Token

There are two ways to store your PAT:

**Option 1:** Using the `credentials` package:

```{r}
#| label: store-pat-1
#| eval: false
# Install credentials if needed
if (!requireNamespace("credentials", quietly = TRUE)) {
  install.packages("credentials")
}

# Replace "YourPAT" with your actual token
credentials::set_github_pat("PositronGitHubPAT")
```

**Option 2:** Store in your `.Renviron` file:

```{r}
#| label: store-pat-2
#| eval: false
# Open .Renviron file
usethis::edit_r_environ()

# In the file that opens, add this line (with your actual token):
# GITHUB_PAT=xxxyyyzzz
# Make sure the file ends with a newline
```

### 4. Restart R

After configuring Git, restart your R session. In Positron you can do this
from the **Console** pane toolbar (the circular-arrow restart icon), from the
Command Palette (`Cmd+Shift+P` / `Ctrl+Shift+P` → "Interpreter: Restart
Interpreter"), or programmatically:

```{r}
#| label: restart
#| eval: false
# Works in both Positron and RStudio
rstudioapi::restartSession()
```

### 5. Verify Settings

Check that everything is configured correctly:

```{r}
#| label: verify
#| eval: false
# Check Git configuration
usethis::git_sitrep()
```

Your output should include:

- Your correct username and email
- Confirmation that your PAT was found (`Personal access token: '<found in env var>'`)

If you see any errors, carefully read the output for troubleshooting guidance.

## Troubleshooting

If you encounter issues:

1. **PAT not found**: Edit your `.Renviron` file manually with
   `usethis::edit_r_environ()` and ensure your token is stored correctly as
   `GITHUB_PAT=your_token_here`.

2. **Git not detected**: Make sure Git is installed and restart Positron.

3. **Authentication failures**: Generate a new PAT if your current one has
   expired.

## Next Steps

Now that you've set up Git and GitHub with R:

1. Create your first Git repository with `usethis::use_git()`
2. Connect it to GitHub with `usethis::use_github()`
3. Learn basic Git commands like commit, push, and pull

## References

- Gist for [Configure GitHub for RStudio](https://gist.github.com/Z3tt/3dab3535007acf108391649766409421)
  (most steps apply identically to Positron)
- Short video playlist:
  [How to Use Git/GitHub with R](https://www.youtube.com/watch?v=k-57UBYnRvw&list=PLSviU861UtD81AuyYb3SbndmAA_qTCoLe)
- [Happy Git with R](https://happygitwithr.com/)
- [usethis package documentation](https://usethis.r-lib.org/)
- [GitHub documentation](https://docs.github.com/)

---

**Congratulations!** You've successfully configured Git and GitHub to work
with Positron. This setup will allow you to use version control for your R
projects and collaborate efficiently with others.
