---
title: "CRISPR target gene annotation map"
author: ""
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
params:
  gene: "nmx"
---

This notebook annotates one CRISPR target gene by walking through three
sequence searches and producing two artifacts: an interactive annotation
map you can drop into a lab report, and a GenBank flat file you can open
in SnapGene or Benchling to design or check edits.

We work on the **padded genomic sequence** of the gene (the gene plus
500 bp of flanking sequence on each side, reverse-complemented if the
gene is on the minus strand). All coordinates in this notebook are
1-based offsets within that padded sequence, not the original chromosome
coordinates.

```{r r-packages}
required_pkgs <- c("gggenomes")
missing_pkgs  <- setdiff(required_pkgs, rownames(installed.packages()))
if (length(missing_pkgs) > 0) {
  install.packages(missing_pkgs, repos = "https://cloud.r-project.org")
}
```

```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(gggenomes)
source("helpers.R")
```

The two chunks above install the one CRAN package the notebook needs
(`gggenomes`, used for the annotation plot) and pull every reusable
function out of `helpers.R`. The helpers contain the FASTA and GFF
parsing, the BLAST and e-PCR wrappers, the annotation table builder,
the plotter, and the GenBank writer.

## Inputs

Per-gene inputs live under `data/<gene_id>/`. Five files are required:

- `genomic.fa`  the padded genomic sequence, one FASTA record.
- `cdna.fa`     the cDNA transcript(s) for this gene.
- `guides.fa`   the sgRNA oligos as FASTA, with `type=CRISPR_guide` in
                each header.
- `primer_pairs.sts`  one line per primer pair in NCBI e-PCR's native
                format (`STS_ID FWD_SEQ REV_SEQ SIZE_RANGE`). The
                STS_ID encodes the cultivar, gene, and both primer
                member names so it carries through to the annotation
                table and the GenBank `/label=` qualifier.
- `gene.gff3`   a subset of the genome's GFF3 containing this gene's
                `gene`, `exon`, and `CDS` records.

```{r paths}
gene_id      <- params$gene
data_dir     <- file.path("data", gene_id)
genomic_path <- file.path(data_dir, "genomic.fa")
cdna_path    <- file.path(data_dir, "cdna.fa")
guides_path  <- file.path(data_dir, "guides.fa")
sts_path     <- file.path(data_dir, "primer_pairs.sts")
gff_path     <- file.path(data_dir, "gene.gff3")

stopifnot(
  "genomic.fa not found"        = file.exists(genomic_path),
  "cdna.fa not found"           = file.exists(cdna_path),
  "guides.fa not found"         = file.exists(guides_path),
  "primer_pairs.sts not found"  = file.exists(sts_path),
  "gene.gff3 not found"         = file.exists(gff_path)
)
```

## Gene

The `load-gene` chunk reads every input file and derives the metadata
the rest of the notebook needs. The `meta` list captures the gene's
chromosome and strand from the GFF3, the gene length, and the offset of
the gene's first base inside the padded genomic sequence (the plot uses
these to draw dashed lines at the gene boundaries).

```{r load-gene}
queries <- parse_fasta(genomic_path)
stopifnot("Expected exactly one record in genomic.fa" = length(queries) == 1)
q <- queries[[1]]
q$gene_name <- gene_id

cdna_records  <- parse_fasta(cdna_path)
guide_records <- parse_fasta(guides_path)
guide_meta    <- parse_oligo_meta(guides_path)
sts_table     <- parse_sts_file(sts_path)
gene_rec      <- read_gene_record(gff_path)

gene_length <- gene_rec$gene_end - gene_rec$gene_start + 1L
padded_len  <- q$length
gene_offset <- (padded_len - gene_length) %/% 2L

meta <- list(
  gene_name    = gene_id,
  chromosome   = gene_rec$chromosome,
  strand       = gene_rec$strand,
  gene_offset  = gene_offset,
  gene_length  = gene_length,
  flank_bp     = gene_offset,
  padded_start = gene_rec$gene_start - gene_offset,
  padded_end   = gene_rec$gene_end   + gene_offset
)

cat(sprintf("Gene:          %s\n", q$gene_name))
cat(sprintf("Locus ID:      %s\n", q$name))
cat(sprintf("Padded length: %d bp\n", q$length))
cat(sprintf("Gene length:   %d bp\n", meta$gene_length))
cat(sprintf("cDNA records:  %d\n", length(cdna_records)))
cat(sprintf("Guides:        %d\n", length(guide_records)))
cat(sprintf("Primer pairs:  %d\n", nrow(sts_table)))
```

Use those counts as your first sanity check before going further. If
the gene length plus 1000 (twice the default flank) does not equal the
padded length, the genomic sequence was clipped at a chromosome edge
and the dashed gene boundaries in later plots will sit closer to the
edges of the figure than the symmetric default.

