## Exons

The genomic sequence we work with includes introns. To draw exons on
the annotation map and to label them in the GenBank file, we ask BLAST
where the cDNA aligns inside the genomic. Each high-scoring pair (HSP)
is one exon.

### Search

`search_cdna()` runs `blastn -task megablast` with the padded genomic
as the query and the cDNA as the subject. The HSP `Q.start` and `Q.end`
columns of the returned data frame are the exon's start and end in the
padded genomic's coordinate system.

```{r blast-exons}
exon_hits <- search_cdna(genomic_path, cdna_path)
cat(sprintf("Exons found: %d HSPs against the cDNA\n", nrow(exon_hits)))
```

### Annotation map

This first plot shows the gene line with the exons drawn as a gold
ribbon below it, plus dashed grey lines at the gene boundaries inside
the padded sequence. It is the same plotting code we will use later
with guides and primers added; right now only exons are populated.

```{r plot-exons, fig.width=10, fig.height=2.6}
annotation_exons <- build_annotation_table(q, meta, exon_hits = exon_hits)
print(plot_annotation(annotation_exons, q, meta))
```

