# Stage 2: cDNA exon annotation

Paste both chunks below into `R/annotation_map.qmd` after the `## Gene`
section. Render to verify, then prompt Claude to commit and push.

## Exon BLAST

Add this section header and chunk:

````markdown
## Exons

```{r blast-exons}
exon_cmd <- sprintf(
  paste(
    'blastn -task megablast -query %s -subject %s',
    '-max_hsps 50',
    '-outfmt "6 qseqid sseqid pident length mismatch gapopen',
    'qstart qend sstart send evalue bitscore qlen slen"'
  ),
  shQuote(genomic_path), shQuote(cdna_path)
)

col_names <- c("Query", "Subject", "Identity", "Aln.Length",
               "Mismatches", "Gaps", "Q.start", "Q.end",
               "S.start", "S.end", "E.value", "Bit.score",
               "Q.len", "S.len")

exon_out <- system(exon_cmd, intern = TRUE)
exon_hits <- if (length(exon_out) > 0) {
  read.table(text = exon_out, sep = "\t", stringsAsFactors = FALSE,
             col.names = col_names)
} else {
  data.frame()
}

cat(sprintf("Exons found: %d HSPs against the cDNA\n", nrow(exon_hits)))
```
````

## Annotation map (gene + exons)

Then add this section and chunk:

````markdown
## Annotation map

```{r plot-exons, fig.width=10, fig.height=2.5}
gene_seq <- data.frame(
  seq_id = q$name,
  length = q$length
)

exon_feats <- if (nrow(exon_hits) > 0) {
  data.frame(
    seq_id    = exon_hits$Query,
    start     = pmin(exon_hits$Q.start, exon_hits$Q.end),
    end       = pmax(exon_hits$Q.start, exon_hits$Q.end),
    feat_type = "Exon",
    stringsAsFactors = FALSE
  )
} else {
  data.frame(seq_id = character(0), start = integer(0), end = integer(0),
             feat_type = character(0), stringsAsFactors = FALSE)
}

p <- gggenomes(seqs = gene_seq, feats = list(exons = exon_feats)) +
  geom_seq() +
  geom_feat(data = feats(exons), aes(color = feat_type),
            position = "identity", linewidth = 5, alpha = 0.8) +
  scale_color_manual(name = NULL, values = c("Exon" = "#DAA520")) +
  labs(title = sprintf("%s (%s)", q$gene_name, q$name)) +
  theme(legend.position = "bottom")

if (!is.null(meta$gene_offset) && meta$gene_offset > 0) {
  gene_start <- meta$gene_offset + 1
  gene_end   <- meta$gene_offset + meta$gene_length
  p <- p +
    geom_vline(xintercept = gene_start, linetype = "dashed",
               color = "grey40", linewidth = 0.4) +
    geom_vline(xintercept = gene_end, linetype = "dashed",
               color = "grey40", linewidth = 0.4)
}

print(p)
```
````
