# Stage 3: Oligo annotation (guides + primer pairs)

Paste the chunks below into `R/annotation_map.qmd` after the `## Exons`
section. Guides are placed with `blastn -task blastn-short`; primer pairs
are placed with e-PCR (`re-PCR -s`) so the amplicon span comes from a
valid pair, not from individual primer hits.

The third chunk replaces the `plot-exons` chunk to add guides and amplicons
to the annotation map.

## Guides via blastn-short

Add this section header and chunk:

````markdown
## Oligos

```{r blast-guides}
guide_cmd <- sprintf(
  paste(
    'blastn -task blastn-short -query %s -subject %s',
    '-word_size 7 -evalue 1000 -dust no -soft_masking false',
    '-ungapped -perc_identity 85 -max_target_seqs 50',
    '-outfmt "6 qseqid sseqid pident length mismatch gapopen',
    'qstart qend sstart send evalue bitscore qlen slen"'
  ),
  shQuote(genomic_path), shQuote(oligo_path)
)

guide_out <- system(guide_cmd, intern = TRUE)
guide_hits <- if (length(guide_out) > 0) {
  read.table(text = guide_out, sep = "\t", stringsAsFactors = FALSE,
             col.names = col_names)
} else {
  data.frame()
}

if (nrow(guide_hits) > 0) {
  guide_hits <- merge(guide_hits, oligo_meta,
                      by.x = "Subject", by.y = "oligo_id", all.x = TRUE)
  guide_hits$strand <- ifelse(guide_hits$S.start <= guide_hits$S.end, "+", "-")
  # Keep guides only, near-full-length, high-identity
  guide_hits <- guide_hits[
    guide_hits$oligo_type == "CRISPR_guide" &
    guide_hits$Aln.Length >= (guide_hits$S.len - 2) &
    guide_hits$Identity >= 95, ]
}

cat(sprintf("Guides placed: %d\n", nrow(guide_hits)))
```
````

## Primer pairs via e-PCR

Add this chunk after the guide chunk:

````markdown
```{r epcr-primers}
# Build the hash that re-PCR queries against. famap turns the FASTA into a
# memory-mapped binary; fahash builds the lookup table on top of it.
famap_file <- tempfile(fileext = ".famap")
hash_file  <- tempfile(fileext = ".hash")
system2("famap",  c("-t", "N", "-b", famap_file, genomic_path),
        stdout = FALSE, stderr = FALSE)
system2("fahash", c("-b", hash_file, famap_file),
        stdout = FALSE, stderr = FALSE)

# Build STS file for re-PCR: one line per primer pair.
oligo_seq_lookup <- setNames(
  vapply(oligo_records, function(x) x$sequence, character(1)),
  vapply(oligo_records, function(x) x$name,     character(1))
)

primer_meta <- oligo_meta[oligo_meta$oligo_type %in%
                          c("forward_primer", "reverse_primer"), ]

pair_ids <- unique(na.omit(primer_meta$pair))
sts_lines <- character(0)
pair_rows <- data.frame(
  pair = character(0), forward_id = character(0), reverse_id = character(0),
  stringsAsFactors = FALSE
)
for (pid in pair_ids) {
  fwd <- primer_meta[primer_meta$pair == pid &
                     primer_meta$oligo_type == "forward_primer", ]
  rev <- primer_meta[primer_meta$pair == pid &
                     primer_meta$oligo_type == "reverse_primer", ]
  if (nrow(fwd) != 1 || nrow(rev) != 1) next
  sts_lines <- c(sts_lines, sprintf("pair_%s\t%s\t%s\t100-10000",
                                     pid,
                                     oligo_seq_lookup[[fwd$oligo_id]],
                                     oligo_seq_lookup[[rev$oligo_id]]))
  pair_rows <- rbind(pair_rows, data.frame(
    pair = pid, forward_id = fwd$oligo_id, reverse_id = rev$oligo_id,
    stringsAsFactors = FALSE
  ))
}

amp_hits <- data.frame(
  seq_id = character(0), pair = character(0), strand = character(0),
  start = integer(0), end = integer(0), size = integer(0),
  forward_id = character(0), reverse_id = character(0),
  stringsAsFactors = FALSE
)

if (length(sts_lines) > 0) {
  sts_file <- tempfile(fileext = ".sts")
  writeLines(sts_lines, sts_file)

  # re-PCR -S takes hash_file then sts_file. Output columns:
  #   sts_name  seq_id  strand  from  to  mism  gaps  act/exp_len
  epcr_out <- system2("re-PCR", c("-S", hash_file, sts_file),
                      stdout = TRUE, stderr = FALSE)
  unlink(sts_file)

  data_lines <- epcr_out[!grepl("^#", epcr_out) & nzchar(epcr_out)]
  if (length(data_lines) > 0) {
    raw <- read.table(text = data_lines, sep = "\t", fill = TRUE,
                      stringsAsFactors = FALSE)
    amp_hits <- data.frame(
      seq_id = raw$V2,
      pair   = sub("^pair_", "", raw$V1),
      strand = raw$V3,
      start  = as.integer(raw$V4),
      end    = as.integer(raw$V5),
      size   = as.integer(raw$V5) - as.integer(raw$V4) + 1L,
      stringsAsFactors = FALSE
    )
    amp_hits <- merge(amp_hits, pair_rows, by = "pair", all.x = TRUE)
  }
}

unlink(c(famap_file, hash_file))

cat(sprintf("Predicted amplicons: %d (from %d pairs in STS)\n",
            nrow(amp_hits), nrow(pair_rows)))
```
````

## Annotation map with guides and amplicons

Replace the `plot-exons` chunk with this expanded version:

````markdown
```{r plot-annotation, fig.width=10, fig.height=3.5}
gene_seq <- data.frame(seq_id = q$name, length = q$length)

exon_feats <- if (nrow(exon_hits) > 0) {
  data.frame(
    seq_id    = exon_hits$Query,
    start     = pmin(exon_hits$Q.start, exon_hits$Q.end),
    end       = pmax(exon_hits$Q.start, exon_hits$Q.end),
    feat_type = "Exon",
    stringsAsFactors = FALSE
  )
} else {
  data.frame(seq_id = character(0), start = integer(0), end = integer(0),
             feat_type = character(0), stringsAsFactors = FALSE)
}

guide_feats <- if (nrow(guide_hits) > 0) {
  data.frame(
    seq_id    = guide_hits$Query,
    start     = pmin(guide_hits$Q.start, guide_hits$Q.end),
    end       = pmax(guide_hits$Q.start, guide_hits$Q.end),
    feat_type = "CRISPR_guide",
    name      = guide_hits$Subject,
    stringsAsFactors = FALSE
  )
} else {
  data.frame(seq_id = character(0), start = integer(0), end = integer(0),
             feat_type = character(0), name = character(0),
             stringsAsFactors = FALSE)
}

# Render amplicons as gggenomes "genes" so the strand arrow is visible.
amp_genes <- if (nrow(amp_hits) > 0) {
  is_rev <- amp_hits$strand == "-"
  data.frame(
    seq_id    = amp_hits$seq_id,
    start     = ifelse(is_rev, amp_hits$end,   amp_hits$start),
    end       = ifelse(is_rev, amp_hits$start, amp_hits$end),
    feat_type = "Amplicon",
    name      = sprintf("pair_%s", amp_hits$pair),
    stringsAsFactors = FALSE
  )
} else {
  data.frame(seq_id = character(0), start = integer(0), end = integer(0),
             feat_type = character(0), name = character(0),
             stringsAsFactors = FALSE)
}

p <- gggenomes(
  seqs  = gene_seq,
  genes = amp_genes,
  feats = list(exons = exon_feats, guides = guide_feats)
) +
  geom_seq() +
  geom_feat(data = feats(exons), aes(y = y - 0.15, yend = y - 0.15,
            color = feat_type),
            position = "identity", linewidth = 5, alpha = 0.8) +
  geom_feat(data = feats(guides), aes(color = feat_type), linewidth = 2) +
  geom_feat_label(data = feats(guides), aes(label = name),
                  size = 2.5, nudge_y = 0.05, color = "#D94040") +
  geom_gene(aes(fill = feat_type), color = NA, size = 10) +
  geom_gene_label(aes(label = name), size = 2.5, nudge_y = 0.05) +
  scale_color_manual(
    name   = NULL,
    values = c("Exon" = "#DAA520", "CRISPR_guide" = "#D94040"),
    labels = c("Exon" = "Exon", "CRISPR_guide" = "CRISPR guide")
  ) +
  scale_fill_manual(
    name   = NULL,
    values = c("Amplicon" = "#40A0D0"),
    labels = c("Amplicon" = "PCR amplicon (e-PCR)")
  ) +
  labs(title = sprintf("%s (%s)", q$gene_name, q$name)) +
  theme(legend.position = "bottom")

if (!is.null(meta$gene_offset) && meta$gene_offset > 0) {
  gene_start <- meta$gene_offset + 1
  gene_end   <- meta$gene_offset + meta$gene_length
  p <- p +
    geom_vline(xintercept = gene_start, linetype = "dashed",
               color = "grey40", linewidth = 0.4) +
    geom_vline(xintercept = gene_end, linetype = "dashed",
               color = "grey40", linewidth = 0.4)
}

print(p)
```
````
