## Oligos

Two kinds of oligos go on the map. **CRISPR guides** are 20mers we
place individually, each one is a single binding site at the edit
target. **Primer pairs** define amplicons: each pair (one forward, one
reverse) produces one PCR product that spans an edit. We treat the two
kinds of oligos differently because the question we ask is different:
"where does this guide bind?" versus "where does this pair amplify?"

### Search

The guide search uses `blastn -task blastn-short`, BLAST's mode for
short queries. It tolerates wider e-value thresholds and word sizes
than default `blastn` so 20mers can be found at all. The helper filters
the raw hits to near-full-length matches at 95% identity or better so
partial or mismatched hits do not clutter the map.

```{r blast-guides}
guide_hits <- search_guides(genomic_path, guides_path)
cat(sprintf("Guides placed: %d\n", nrow(guide_hits)))
```

The primer search uses e-PCR. Unlike BLAST, e-PCR asks "does this
forward primer paired with this reverse primer produce a product, and
if so, where?" The `search_primer_pairs()` helper builds the e-PCR
index (`famap` then `fahash`) on the padded genomic, runs `re-PCR -S`
against the gene's `primer_pairs.sts`, and decodes each STS_ID into
the matching forward and reverse oligo IDs so the annotation table can
label each primer arrow with its real name.

```{r epcr-primers}
amp_hits <- search_primer_pairs(genomic_path, sts_path, sts_table, gene_id)
cat(sprintf("Predicted amplicons: %d (from %d pairs in STS)\n",
            nrow(amp_hits), nrow(sts_table)))
```

### Annotation table

We now have three search outputs: exons, guides, and amplicons.
Instead of letting each downstream step (plot, GenBank file) construct
its own per-feature data structures from these, we build one **unified
annotation table** that lists every feature in one place. The plot and
the GenBank writer both consume this table, so the two artifacts can
never disagree about what was found.

```{r build-annotation-table}
annotation <- build_annotation_table(q, meta, exon_hits, guide_hits, amp_hits)
cat(sprintf("Annotation rows: %d\n", nrow(annotation)))
knitr::kable(
  with(annotation, data.frame(
    feat_type, name, start, end, strand,
    note = ifelse(nchar(note) > 40, paste0(substr(note, 1, 37), "..."), note)
  )),
  caption = "Unified annotation table"
)
```

Each row is one annotation. The amplicon's `name` is the STS_ID, the
primer rows derive their positions from the amplicon edges, and the
gene + exon rows reuse the GFF coordinates. `feat_type` drives both
the visual style on the plot and the GenBank feature type on export.

### Annotation map

The plot reads only the annotation table. Exons sit below the gene
line as a gold ribbon, guides as red bars on the line, the amplicon as
a thin blue span above it, and the forward (green) and reverse
(purple) primer arrows at the amplicon edges. Dashed grey lines mark
the gene boundaries inside the padded sequence.

```{r plot-annotation, fig.width=10, fig.height=2.6}
print(plot_annotation(annotation, q, meta))
```

